ChemKED: a human- and machine-readable data standard for chemical kinetics experiments

### 2017-02-22 pyked.tex
incomplete additions on related work

\usepackage{todonotes}
\newcommand\ck{ChemKED}
\usepackage[inline]{enumitem}
\setlist{noitemsep}
%======================================================================
\addbibresource{pyked.bib}
%======================================================================
\newcommand\papertopic{Reaction Kinetics}
%======================================================================

@@

@@

The \ck{} file is a YAML-format file. This format offers the advantages of being
human readable, written in plain text, and having parsers in most common
programming languages, including Python, C++, Java, Perl, MATLAB, and many
others. The YAML syntax is quite simple. The basic file structure is made of
mappings, delimited by a colon. The key for the mapping is on the left of the
colon, and the value on the right can be a single value, a sequence of values,
a nested mapping, or some combination of these:

\begin{minted}{yaml}
    key1: value  # Single-value mapping
    key2: [0, 1, 2]  # Mapping with sequence
    key3:  # Alternative sequence format
      - value
      - value
    key4:  # Nested mapping
      key5: 0
\end{minted}

The \yaml|value| can be a string, integer, or floating point number. The type is
inferred from the \yaml|key| when the value is read. The \ck{} format is
designed to include all of the information necessary to simulate a given
experiment and can be broken into two main sections. The first section details
all of the ``meta'' information about the experiment and the \ck{} file. In this
section, the common \yaml|author| mapping is used for any fields where an author
must be specified. The \yaml|author| mapping contains the following fields:
%
\begin{itemize}
    \item \yaml|name| (required, string): The author's full name
    \item \yaml|ORCID| (optional, string): the author's ORCID number
\end{itemize}
%
More generally for the meta information, the required keys in this section
are:
%
\begin{itemize}
    \item \yaml|file-author| (required, \yaml|author|-type mapping): The author
    of the \ck{} file
    \item \yaml|file-version| (required, integer): The version of the \ck{} file
    \item \yaml|chemked-version| (required, string): The version of the \ck{}
    schema that this file conforms to
    \item \yaml|reference| (required, mapping): The reference information for
    the article associated with the data in the file. The fields in this
    mapping are:
    \begin{itemize}
        \item \yaml|authors| (required, sequence): A sequence of \yaml|author|
        mappings
        \item \yaml|journal| (required, string): The name of the publishing
        journal
        \item \yaml|year| (required, integer): The year of publication
    \end{itemize}
    \item \yaml|experiment-type| (required, string): One of \yaml|ignition
    delay| or \yaml|flame speed|
    \item \yaml|apparatus| (required, string): The type of apparatus used to
    perform the experiment
\end{itemize}

@@

>>>>

\usepackage{todonotes}
\newcommand\ck{ChemKED}
\newcommand\pk{\texttt{PyKED}}
\usepackage[inline]{enumitem}
\setlist{noitemsep}
%======================================================================
\addbibresource{pyked.bib}
%======================================================================
\newcommand\papertopic{Reaction Kinetics}
%======================================================================

@@

The combustion\slash chemical kinetics community lacks an accepted, commonly used
standard for recording and sharing data from fundamental combustion experiments.
Interestingly, this contrasts the situation for chemical kinetic models, where the
CHEMKIN~\cite{Kee:1996ck} format dominates. Competing standards such as the
Cantera~\cite{Cantera:2.3.0} CTI format or FlameMaster lag behind considerably, although
use of the former continually grows due to Cantera's open availability.\todo{add flamemaster citation}{}
Similarly,

Instead, such data must typically be found in under-documented, nonstandard comma-separated value
files and Excel spreadsheets, or even contained in PDF tables or figures rather than as
supplementary material associated with a paper.

While the community has not widely adopted a standard format,
---most notably, Frenklach developed the PrIMe (\textbf{Pr}ocess \textbf{I}nformatics
\textbf{M}od\textbf{e}l) data format~\cite{Frenklach:2007bm,You:2011hy} and an associated
online database, the PrIMe Data Warehouse~\cite{PrIMe}.

More recently, Varga et al.~\cite{Varga2015a,Varga2015b} developed the \texttt{ReSpecTh} standard,
an XML-based format that builds on, and is compatible with, PrIMe.

We aim to develop a new, human- and machine-readable data standard for fundamental combustion
experiments, \ck{}, and offer tools for easily working with data encoded in this format.
Our motivations are similar to some of those of the PrIMe and \texttt{ReSpecTh} teams, although
less ambitious in scope: enable easy sharing and use of data, for the (primary) purposes of
developing and validating predictive chemical kinetic models.
However, those efforts were generally less successful in being adopted---we hope to succeed
by developing more user-friendly data files and tools for interacting with them.

Niemeyer recently introduced an initial version of the \ck{} format~\cite{Niemeyer:2016wf},
and in this work we further formalize and develop the standard. We also discuss a related
Python software tool for validating and working with \ck{} files: \pk{}.

@@

The \ck{} file is a YAML-format file. This format offers the advantage of being
human readable, written in plain text, and having parsers in most common
programming languages, including Python, C++, Java, Perl, MATLAB, and many
others. The YAML syntax is quite simple. The basic file structure is made of
mappings, delimited by a colon. The key for the mapping is on the left of the
colon, and the value on the right can be a single value, a sequence of values, or a nested mapping:

\begin{minted}{yaml}
    key1: value  # Single-value mapping
    key2: [0, 1, 2]  # Mapping with sequence
    key3:  # Alternative sequence format
      - 0
      - 1
    key4:  # Nested mapping
      key5: 0
\end{minted}

The \yaml|value| can be a string, integer, or floating point number. The type is
inferred from the \yaml|key| when the value is read. The \ck{} format is
designed to include all of the information necessary to simulate a given
experiment. The \ck{} format can be broken into two main sections. The first
section details all of the ``meta'' information about the experiment and the
\ck{} file. In this section, the common \yaml|author| mapping is used for any
fields where an author must be specified. The \yaml|author| mapping contains the
following fields:
%
\begin{itemize}
    \item \yaml|name| (required, string): The author's full name
    \item \yaml|ORCID| (optional, string): the author's ORCID number
\end{itemize}
%
More generally for the meta information, the required keys in this section
are:
%
\begin{itemize}
    \item \yaml|file-author| (required, \yaml|author|-type mapping): The author
    of the \ck{} file
    \item \yaml|file-version| (required, integer): The version of the \ck{} file
    \item \yaml|chemked-version| (required, string): The version of the \ck{}
    schema that this file conforms to
    \item \yaml|reference| (required, mapping): The reference information for
    the article associated with the data in the file. The fields in this
    mapping are:
    \begin{itemize}
        \item \yaml|authors| (required, sequence): A sequence of \yaml|author| mappings
        \item \yaml|journal| (required, string): The name of the publishing journal
        \item \yaml|year| (required, integer): The year of publication
    \end{itemize}
    \item \yaml|experiment-type| (required, string): One of \yaml|ignition delay| or \yaml|flame speed|
    \item \yaml|apparatus| (required, string): The type of apparatus used to perform the experiment
\end{itemize}

@@

\pk{} relies on well-established scientific Python software tools.
These include NumPy~\cite{vanderWalt:2011np} for array manipulation, Pint~\cite{Grecco2016} for
interpreting and converting between units, \texttt{ruamel.yaml} for parsing YAML files,\todo{needs ref}{}
Cerberus for validating \ck{} files\todo{needs ref}{},
and pytest~\cite{pytest:3.0.1} for unit testing.
Travis-CI~\cite{Travis2016} also provides continuous integration testing.

\todo[inline]{Need to archive version of PyKED and cite}

\pk{} is available under the permissive, open-source BSD 3-clause license via a GitHub
repository: \url{https://github.com/pr-omethe-us/PyKED}. It can be installed using setuptools
by downloading the source code files and executing \texttt{python setup.py install}. More mature
versions of \pk{} will be distributed on \href{https://pypi.python.org/pypi}{PyPI}
(Python Package Index) and conda-forge.\todo{add URL}{}
